---
title: "Setting up RStudio inside a container"
author: "Peter Smith"
date: "11 May 2020"
output:
  html_document:
    toc: yes
    toc_collapsed: yes
    toc_float: yes
---




#Ubuntu packages to install

Basically, in RStudio, we want to get through running these commands without issue to say that everythings installed correctly.
```{r, eval=FALSE}
install.packages('tidyverse')
install.packages('odbc')
```

This will highlight most Linux-packages that are missing and need to be installed. You will get a bunch of error messages of the following form.

```{r, eval=FALSE}
Using PKG_CFLAGS=
Using PKG_LIBS=-lxml2
------------------------- ANTICONF ERROR ---------------------------
Configuration failed because libxml-2.0 was not found. Try installing:
 * deb: libxml2-dev (Debian, Ubuntu, etc)
 * rpm: libxml2-devel (Fedora, CentOS, RHEL)
 * csw: libxml2_dev (Solaris)
If libxml-2.0 is already installed, check that 'pkg-config' is in your
PATH and PKG_CONFIG_PATH contains a libxml-2.0.pc file. If pkg-config
is unavailable you can set INCLUDE_DIR and LIB_DIR manually via:
R CMD INSTALL --configure-vars='INCLUDE_DIR=... LIB_DIR=...'
```

just install whatever Ubuntu package it's looking for (in the above case libxml2-dev. In the end I needed to run the following commands

```{r, eval=FALSE}
#for tidyverse
sudo apt-get install libssl-dev
sudo apt-get install libcurl4-openssl-dev
sudo apt-get install libxml2-dev

#for odbc
sudo apt-get install unixodbc-dev
```

for the odbc package I had an issue where I had to remove and reinstall it. I don't think future users would have this issue though, but if they do, removing and reinstalling follows the same approach as the Stringi issue below.

#Stringi issue

There is also some issue with stringi that may need to be sorted out.

The error I got was while trying to install tidyverse
```{r, eval=FALSE}
Error in dyn.load(file, DLLpath = DLLpath, ...) : 
  unable to load shared object '/home/peter/R/x86_64-pc-linux-gnu-library/3.6/stringi/libs/stringi.so':
  libicui18n.so.60: cannot open shared object file: No such file or directory
Calls: <Anonymous> ... namespaceImport -> loadNamespace -> library.dynam -> dyn.load
Execution halted
```

googling this leads to these links, with the advice given to just delete and reinstall stringi.

```{r, eval=FALSE}
https://community.rstudio.com/t/unable-to-load-packages-cannot-open-shared-object-file/47453
https://forum.manjaro.org/t/manjaro-update-now-r-cant-load-libraries/115419/2
```

*Delete stringi

delete stringi (or just move it in the case below since I was nervous)

```{r, eval=FALSE}
cd ~/R/x86_64-pc-linux-gnu-library/3.6
mkdir ~/test
mv stringi/* ~/test/
rmdir stringi
```

Then reinstall in R-console

```{r, eval=FALSE}
install.packages("stringi")
install.packages('tidyverse')
```

tidyverse now installed without issue.

#Vertica connection

##Vertica Driver
No one else should need to do this, but I uploaded and extracted vertica drivers from Vertica website onto root. The following driver should exist, otherwise it needs to be installed

```{r, eval=FALSE}
/opt/vertica/lib64/libverticaodbc.so
```

If it doesn't exist for whatever reason. find drivers from vertica, and upload them.
to unzip, I used,
```{r, eval=FALSE}
tar -xzf vertica-client-9.3.1-0.x86_64.tar.gz
```
then moved it to the root folder so everyone would have access to it. 

##Create .odbc.ini file 
the .odbc.ini file needs to be created in your home directory. This file stores connection details that R will use when connecting to any database.

To create your first connection, first navigate to your homefolder

```{r, eval=FALSE}
cd ~
```

then use vim to create a connection
```{r, eval=FALSE}
vim .odbc.ini
```

For example, my .odbc.ini folder has the following entry:

```{r, eval=FALSE}
[vertica]
Driver=/opt/vertica/lib64/libverticaodbc.so
Server=10.0.0.211
Database=
Port=5433
ResultBufferSize=0
UID=xxxxx
Password=xxxxx
```

NOTE: The username and password here are NOT your usual account username and passwords, get these from whoever maintains the SQL database you want to connect to. 


At this point your home directory should look something like this:

```{R, eval=FALSE}
peter@sb-rstudio:~$ ls -lhart
total 36M
-rw-r--r--  1 peter 2157  807 Apr 29 12:11 .profile
-rw-r--r--  1 peter 2157 3.7K Apr 29 12:11 .bashrc
-rw-r--r--  1 peter 2157  220 Apr 29 12:11 .bash_logout
drwxr-xr-x  5 root  root 4.0K Apr 29 12:11 ..
drwx------  2 peter 2157 4.0K Apr 29 12:11 .cache
drwxr-xr-x  3 peter 2157 4.0K Apr 29 12:11 R
drwxr-xr-x  3 peter 2157 4.0K May  5 18:44 .config
drwx------  3 peter 2157 4.0K May  5 19:15 .gnupg
-rw-r--r--  1 peter 2157 1.3K May 11 20:27 .Rhistory
drwxr-xr-x 15 peter 2157 4.0K May 11 20:27 .rstudio
-rw-------  1 peter 2157 1.1K May 11 20:27 .bash_history
-rw-r--r--  1 peter 2157  151 May 11 20:38 .odbc.ini
-rw-------  1 peter 2157 2.7K May 11 20:38 .viminfo
drwxr-xr-x  9 peter 2157 4.0K May 11 20:38 .
```

##R-Libraries and creating connection

First make sure that the necessary packages are installed
```{r, eval=FALSE}
library(DBI)
library(odbc)

#if the above packages aren't found, then run:
  #install.packages('DBI')
  #install.packages('odbc')

con <- dbConnect(odbc::odbc(), "vertica", bigint="integer")

```

This connection should just work. You can validate this by running a query. If you're struggling with the above, using the New Connection tab on one of your R-panes can also help to check what connections is can detect.



That's it!


#Roughwork

YOU PROBABLY DON'T NEED TO READ THIS. THIS WAS MY NOTES AS I WAS RECORDING THEM STEP BY STEP - BASICALLY AN UNNEAT VERSION OF THE ABOVE


-------------------------------------
--New container:
-------------------------------------


basically, installing tidyverse highlights a bunch of things I need to install with sudo.

1) in R, tried install.packages('tidyverse')
get a whole bunch of errors of the form


```{r, eval=FALSE}
Using PKG_CFLAGS=
Using PKG_LIBS=-lxml2
------------------------- ANTICONF ERROR ---------------------------
Configuration failed because libxml-2.0 was not found. Try installing:
 * deb: libxml2-dev (Debian, Ubuntu, etc)
 * rpm: libxml2-devel (Fedora, CentOS, RHEL)
 * csw: libxml2_dev (Solaris)
If libxml-2.0 is already installed, check that 'pkg-config' is in your
PATH and PKG_CONFIG_PATH contains a libxml-2.0.pc file. If pkg-config
is unavailable you can set INCLUDE_DIR and LIB_DIR manually via:
R CMD INSTALL --configure-vars='INCLUDE_DIR=... LIB_DIR=...'
--------------------------------------------------------------------
```


So I try installing each one in terminal, then retry installing R until I've installed all the necessary stuff.
in terminal: 

```{r, eval=FALSE}
sudo apt-get install libssl-dev
sudo apt-get install libcurl4-openssl-dev
sudo apt-get install libxml2-dev

```

after this, installing tidyverse gives the following error

```{r, eval=FALSE}
Error in dyn.load(file, DLLpath = DLLpath, ...) : 
  unable to load shared object '/home/peter/R/x86_64-pc-linux-gnu-library/3.6/stringi/libs/stringi.so':
  libicui18n.so.60: cannot open shared object file: No such file or directory
Calls: <Anonymous> ... namespaceImport -> loadNamespace -> library.dynam -> dyn.load
Execution halted
```

googling this leads to this link

```{r, eval=FALSE}
https://community.rstudio.com/t/unable-to-load-packages-cannot-open-shared-object-file/47453
https://forum.manjaro.org/t/manjaro-update-now-r-cant-load-libraries/115419/2
```

from there, there is the advice:

```{r, eval=FALSE}
You have to rebuild the stringi library. Not as root, since other R libraries depend on the user stringi.so library (the path is important in this case).

The easiest way might be simply removing the /home/chance/R/x86_64-pc-linux-gnu-library/3.6/stringi folder and "reinstall" stringi as a user . Which means build it against the new icu library.
```


I'm a bit nervoes to follow instructions, but here goes


delete stringi
```{r, eval=FALSE}
cd ~/R/x86_64-pc-linux-gnu-library/3.6
mkdir ~/test
mv stringi/* ~/test/
rmdir stringi
```

in console

```{r, eval=FALSE}
install.packages("stringi")
install.packages('tidyverse')
```

and voila - tidyverse is installed. as a note, appare

OK, let's see if I can get that connection to vertica working


##connecting to vertica

OK, so first, I downloaded drivers from Vertica's website and uploaded it onto the Rstudio container

then I unzipped it in my home directory with 
```{r, eval=FALSE}
tar -xzf vertica-client-9.3.1-0.x86_64.tar.gz
```
I then moved the opt folder this creates into root so everyone can access it


ok error time. try installing packages

```{r, eval=FALSE}
library(odbc)
Error: package or namespace load failed for ???odbc??? in dyn.load(file, DLLpath = DLLpath, ...):
 unable to load shared object '/home/peter/R/x86_64-pc-linux-gnu-library/3.6/odbc/libs/odbc.so':
  libiodbc.so.2: cannot open shared object file: No such file or directory
```

ok I think I had tried installing odbc already. let me delete it first and start from scratch

in console
```{r, eval=FALSE}
peter@sb-rstudio:~/R/x86_64-pc-linux-gnu-library/3.6$ rm -r odbc
```

then we get familiar errors when trying to reinstall odbc

```{r, eval=FALSE}
sudo apt-get install unixodbc-dev
```

now install.packages('odbc') installs fine
if running it gives a corrupt file issue, restarting R with  .rs.restartR() fixed the issue for me


```{r, eval=FALSE}
peter@sb-rstudio:~$ ls
R  test  vertica-client-9.3.1-0.x86_64.tar.gz
```

now, I created a .odbc.ini folder in my home directory.

my home directory currently looks as follows

```{R, eval=FALSE}
peter@sb-rstudio:~$ ls -lhart
total 36M
drwxr-xr-x  3 peter 2157 4.0K Jan  3 00:38 opt
-rw-r--r--  1 peter 2157  807 Apr 29 12:11 .profile
-rw-r--r--  1 peter 2157 3.7K Apr 29 12:11 .bashrc
-rw-r--r--  1 peter 2157  220 Apr 29 12:11 .bash_logout
drwxr-xr-x  5 root  root 4.0K Apr 29 12:11 ..
drwx------  2 peter 2157 4.0K Apr 29 12:11 .cache
drwxr-xr-x  3 peter 2157 4.0K Apr 29 12:11 R
drwxr-xr-x  3 peter 2157 4.0K May  5 18:44 .config
drwx------  3 peter 2157 4.0K May  5 19:15 .gnupg
-rw-r--r--  1 peter 2157  36M May  7 10:13 vertica-client-9.3.1-0.x86_64.tar.gz
drwxr-xr-x  8 peter 2157 4.0K May 11 20:03 test
-rw-r--r--  1 peter 2157 1.3K May 11 20:27 .Rhistory
drwxr-xr-x 15 peter 2157 4.0K May 11 20:27 .rstudio
-rw-------  1 peter 2157 1.1K May 11 20:27 .bash_history
-rw-r--r--  1 peter 2157  151 May 11 20:38 .odbc.ini
-rw-------  1 peter 2157 2.7K May 11 20:38 .viminfo
drwxr-xr-x  9 peter 2157 4.0K May 11 20:38 .
```

note the opt folder and the .odbc.ini file. These are the 2 important things. inside the .odbc I have the following connection details

```{r, eval=FALSE}
[vertica]
Driver=/opt/vertica/lib64/libverticaodbc.so
Server=10.0.0.211
Database=
Port=5433
ResultBufferSize=0
UID=xxxxx
Password=xxxxx
```


